|

### 2023-09-20 paper.Rmd
background done

Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook1999applied. Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns. There can be various types of departures from this ideal pattern. Non-linearity, heteroskedasticity and non-normality, shown in Figure \ref{fig:residual-plot-common-departures}B,C,D are three commonly checked departures.

@@

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper. (@loy2016variations discuss related work on non-normality checking.)

@@

The effectiveness of lineup protocol for regression analysis has been validated by @majumder2013validation under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a corresponding conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where the contamination of the data violate the assumptions of the convetional test, visual test outperforms the conventional test by a large margin. This is encouraging, as it supports the use of visual inference in situations where there are no existing numerical testing procedures. Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals [@loy2013diagnostic, @loy2014hlmdiag and @loy2015you].

>>>>

Graphical summaries where residuals are plotted against fitted values, or other functions of the predictors (expected to be approximately orthogonal to the residuals) are considered to be the most important residual plots by @cook1999applied. Figure \ref{fig:residual-plot-common-departures}A shows an example of an ideal residual plot where points are symmetrically distributed around the horizontal zero line (red), with no discernible patterns. There can be various types of departures from this ideal pattern. Non-linearity, heteroskedasticity and non-normality, shown in Figures \ref{fig:residual-plot-common-departures}B, \ref{fig:residual-plot-common-departures}C, and \ref{fig:residual-plot-common-departures}D are three commonly checked departures.

@@

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot where the residuals have a skewed distribution, as seen by the uneven vertical spread. Unlike non-linearity and heteroskedasticity, non-normality is usually detected with a different type of residual plot: a histogram or a normal probability plot.  Because we focus on scatterplots, non-normality is not one of the departures examined in depth in this paper. (@loy2016variations discuss related work on non-normality checking.)

@@

The effectiveness of lineup protocol for regression analysis has been validated by @majumder2013validation under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power to a t-test, though they express that in general it is unnecessary to use visual inference if there exists a corresponding conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where the contamination of the data violate the assumptions of the convetional test, visual test outperforms the conventional test by a large margin. This is encouraging, as it supports the use of visual inference in situations where there are no existing numerical testing procedures. Visual inference has also been integrated into diagnostics for hierarchical linear models where the lineup protocol is used to judge the assumptions of linearity, normality and constant error variance for both the level-1 and level-2 residuals \citep{loy2013diagnostic, loy2014hlmdiag, loy2015you}.
